Forecasting the US Election 2024

### 2024-11-04 paper/paper.qmd
Edit writing for abstract and appendix

thanks: "Code and data are available at: [https://github.com/aj3616/Forcasting-US-Elections](https://github.com/aj3616/Forcasting-US-Elections)."
date: today
date-format: long
abstract: "This paper uses polling data to project the outcome of the 2024 U.S. Presidential election between Kamala Harris and Donald Trump. By employing a multiple linear regression model, we predict each candidate’s level of support across states and estimate the corresponding electoral vote counts. Our analysis suggests that Harris has a potential advantage in the Electoral College, highlighting the influence of polling trends and pollster-specific biases on election forecasting. These findings underscore the importance of aggregating diverse polling data to create a more balanced and accurate prediction of election results."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

As of October-November 2024, the U.S. presidential election is almost at its final stage, with all candidates actively campaigning across numerous states. The presidential election will happen on November 5, 2024 [@calendar]. The two major parties, Democrats and Republicans, are in particular focus. Former President Donald Trump represents the Republican party, and President Joe Biden initially led re-election for the Democrats. However, Biden has dropped out, and Kamala Harris has taken over as the Democratic nominee [@g24], who is the first Black woman nominee. Harris is the Both candidates are focusing their efforts on securing votes by addressing their aspect on social issues.

Pollsters and election forecasting with data evidence are important in predicting election outcomes. Polling can highlight underlying or emerging issues, as well as reveal voter preferences. The polling outcomes allow campaigns of different parties to strategically target specific demographics. For example in 2012, pollsters and data modellers such as Nate Silver have all used survey research and statistical models to successfully predict the result of Barack Obama's victory in the presidential election with consistent and reliable forecasts [@b17].

The rest of this paper is organized as follows. @sec-data covers the data utilized in this analysis, highlighting key variables and sources, particularly focusing on the quality metrics that influence polling accuracy. @sec-model details our modeling approach for each candidate, drawing on insights from recent electoral cycles. @sec-results provides our predictions for the Electoral College based on the model outputs.  @sec-discussion examines the implications of our findings and offers suggestions for future research. Finally, the @sec-appendixa and @sec-appendixb assesses YouGov's polling methodology alongside our proposed survey methodology.

@@

We use the polling data to forecast the potential outcomes of the 2024 U.S. presidential election between Kamala Harris and Donald Trump.The dataset was obtained from FiveThirtyEight’s “Poll of Polls” for the 2024 U.S. Presidential election [@fivethirtyeight2024]. It provides a comprehensive view of voter preferences through aggregated results from numerous national polls conducted by various polling organizations.

@@

YouGov has been a popular polling organization in predicting U.S. elections for a long time, they have experience predicting US elections in the past. One of the key reasons for us to select YouGov to dive into it is their innovative use of an advanced statistical model known as Multilevel Regression with Post-stratification (MRP). This model allows YouGov to have detailed predictions by using survey data with demographic background information, ensuring the representativeness of broader populations.

Moreover, FiveThirtyEight found that YouGov did not exhibit a strong house effect during the 2016 U.S. election, meaning they did not show a bias toward any political party[@n16], this further built up their credibility. Their model is intended to predict the votes of everyone in the US national voter file. YouGov’s MRP model used to predict support for each candidate in the US 2024 presidential election has three parts. Firstly, they will use people’s responses in the survey to estimate the likelihood of voting. From here, calculate their probability of voting for a specific party if people will vote. Combined and reweight responses, to give predictions for candidates [@br24].

@@

YouGov's frame of the sample is people in the YouGov panel which use a nonprobability sampling. Everyone can join the YouGov panel, and YouGov will recruit participants (the sample) to the survey with a rigours process. YouGov recruits American adults through various advertising methods and partnerships. They also offer surveys in multiple languages.

@@

Phone Outreach - Random Digit Dialing (RDD): We will employ a random-digit dialling method to recruit respondents for phone interviews. This approach allows us to reach individuals who may not be engaged online, particularly older demographics and those in rural areas. However, this traditional recruitment format will constitute only a small portion of our overall efforts, as advancements in technology have made online recruitment significantly more efficient. We plan to use phone outreach as a coplelement when online recruitment does not yield sufficient participants. This approach helps avoid inclusion errors by reaching populations without internet access or those who are not engaged in online communities. However, we face challenges such as increasing non-response rates due to the public's growing reluctance to participate in phone surveys and the rising use of technologies that screen unsolicited calls [@wang2015forecasting].

>>>>

thanks: "Code and data are available at: [https://github.com/aj3616/Forcasting-US-Elections](https://github.com/aj3616/Forcasting-US-Elections)."
date: today
date-format: long
abstract: "This paper uses polling data to project the outcome of the 2024 U.S. Presidential election between Kamala Harris and Donald Trump. By employing a multiple linear regression model, we predict each candidate’s level of support across states and estimate the corresponding electoral vote counts. Our analysis suggests that Harris has a potential advantage in the Electoral College, winning 48.1% support, highlighting the influence of polling trends and pollster-specific biases on election forecasting. These findings underscore the importance of aggregating diverse polling data to create a more balanced and accurate prediction of election results."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

As of October-November 2024, the U.S. presidential election is almost at its final stage, with all candidates actively campaigning across numerous states. The presidential election will happen on November 5, 2024 [@calendar]. The two major parties, Democrats and Republicans, are in particular focus. Former President Donald Trump represents the Republican party, and President Joe Biden initially led re-election for the Democrats. However, Biden dropped out, and Kamala Harris took over as the Democratic nominee [@g24], the first Black woman nominee. Both candidates are focusing their efforts on securing votes by addressing their aspect on social issues.

Pollsters and election forecasting with data evidence are important in predicting election outcomes. Polling can highlight underlying or emerging issues, as well as reveal voter preferences. The polling outcomes allow campaigns of different parties to strategically target specific demographics. For example in 2012, pollsters and data modellers such as Nate Silver have all used survey research and statistical models to successfully predict the result of Barack Obama's victory in the presidential election with consistent and reliable forecasts [@b17]. In our study, we build on this tradition by leveraging polling data to forecast the outcome of the 2024 U.S. Presidential election. Using a multiple linear regression model, we predict each candidate’s support across various states. This approach not only provides an indication of the potential Electoral College outcome but also illustrates the impact of polling trends and pollster-specific biases on election forecasting. By aggregating data from multiple sources, our analysis underscores the value of comprehensive polling and enhancing the accuracy of election predictions.

The rest of this paper is organized as follows. @sec-data covers the data utilized in this analysis, highlighting key variables and sources, particularly focusing on the quality metrics that influence polling accuracy. @sec-model details our modelling approach for each candidate, drawing on insights from recent electoral cycles. @sec-results provides our predictions for the Electoral College based on the model outputs.  @sec-discussion examines the implications of our findings and offers suggestions for future research. Finally, the @sec-appendixa and @sec-appendixb assesses YouGov's polling methodology alongside our proposed survey methodology.

@@

We use the polling data to forecast the potential outcomes of the 2024 U.S. presidential election between Kamala Harris and Donald Trump. The dataset was obtained from FiveThirtyEight’s “Poll of Polls” for the 2024 U.S. Presidential election [@fivethirtyeight2024]. It provides a comprehensive view of voter preferences through aggregated results from numerous national polls conducted by various polling organizations.

@@

YouGov has been a popular polling organization in predicting U.S. elections for a long time, they have experience predicting US elections in the past. One of the key reasons for us to select YouGov to dive into it is their innovative use of an advanced statistical model known as Multilevel Regression with Post-stratification (MRP). This model allows YouGov to have detailed predictions by using survey data with demographic background information, ensuring the representativeness of broader populations; and accurately estimating voter preferences across diverse groups.

Moreover, FiveThirtyEight found that YouGov did not exhibit a strong house effect during the 2016 U.S. election, meaning they did not show a bias toward any political party[@n16], this further built up their credibility. Their model is intended to predict the votes of everyone in the US national voter file. YouGov’s MRP model used to predict support for each candidate in the US 2024 presidential election has three parts. Firstly, they will use people’s responses in the survey to estimate the likelihood of voting. From here, calculate their probability of voting for a specific party if people will vote. Combined and reweighted responses, to give predictions for candidates [@br24].

@@

YouGov's frame of the sample is people in the YouGov panel which uses a nonprobability sampling. This means not every individual in the population has an equal chance of being chosen; samples are selected based on factors other than randomness [@vehovar2016non]. Everyone can join the YouGov panel, and YouGov will recruit participants (the sample) to the survey with a rigours process. YouGov recruits American adults through various advertising methods and partnerships. They also offer surveys in multiple languages.

@@

Phone Outreach - Random Digit Dialing (RDD): We will employ a random-digit dialling method to recruit respondents for phone interviews. This approach allows us to reach individuals who may not be engaged online, particularly older demographics and those in rural areas. However, this traditional recruitment format will constitute only a small portion of our overall efforts, as advancements in technology have made online recruitment significantly more efficient. We plan to use phone outreach as a complement when online recruitment does not yield sufficient participants. This approach helps avoid inclusion errors by reaching populations without internet access or those who are not engaged in online communities. However, we face challenges such as increasing non-response rates due to the public's growing reluctance to participate in phone surveys and the rising use of technologies that screen unsolicited calls [@wang2015forecasting].
